\mytitle{}

### 2018-11-07 paper/paper.tex
Draft 1...

\emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).

?? We explore the design space, and improve upon existing approaches.

We present...
?? Contribs

?? Taking up space to figure out how much room I have for an intro

?? still taking up space...

?? still going...

?? done...
\end{abstract}

@@

To date, there have been few applications of this class of algorithms towards intrusion prevention which make use of their full potential for online control, rather than using them as the basis of a classifier.
We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
In service of this goal, this paper contributes:
\begin{itemize}
	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:rethinking-the-state-space}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws, and risk factors concerning many traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item A source-level granularity approach to RL-driven DDoS prevention, improving upon past aggregate-based approaches (\cref{sec:rethinking-the-state-space}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

\section{Acting On Individual Sources}
Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per source.
However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our enhancements to the RL model of \textcite{DBLP:journals/eaai/MalialisK15}, present a new model which draws on domain knowledge leveraged by SPIFFY \cite{DBLP:conf/ndss/KangGS16}, and offer an empirical investigation into the effectiveness of flow-level statistics on live decision-making.

@@

%?? Why does SPF only do well sometimes? Model is actually more difficult to learn, so it seems to do best when it has a larger set of decisions to learn from. But, it does worse for TCP?
\fakepara{Model performance}
Of the results presented, SPF's meandering performance is interesting given its far smaller action space.
It's natural to expect that this would the model easier to learn, but the additional state required appears to make the model \emph{harder}, beyond even the gains we might expect from choosing a non-zero discount factor.
Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
Paradoxically, Marl++ generally achieves the best performance yet actively suffers when treated as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.

@@

\section{Future Work}

Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.

?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.

?? IDEA: try out average reward, TD($\lambda$) methods as future work...

?? What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge? I think a good candidate is the sum of up and down throughputs (normalised by capacity sum), so long as \emph{neither exceeds the link capacity}. We can extend the team-based formulations similarly. This, in theory, promotes traffic diversity since it's not like flooding-based DDoS attacks are going to submit meaningful work to a server. The intuition, I suppose, is that certain classes of flow will have a small footprint in one direction which causes a sizeable increase in the other! Alternatively, monitor the health of canary flows which cross the team boundary (i.e. only one in-out link).

?? IDEA: Apply these techniques to programmable data planes etc. While it's pretty neat that what we have works assuming that ach router is a software (x86) switch running OVS, what might we need to consider when applying this to `real' switches? ``PDP can allow this to be added to real routers to make it efficient to keep \& process state in the manner we require, as well as enabling more adaptive deployment''. Cite P4, BPFabric, other work on PDPs?

?? Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e. viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance. It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques. THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?

?? What concessions will we have to make in order to make per-flow processing more viable? Intelligent sampling/reanalysis of flows when needed (i.e. an external heuristic guiding method)? In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!

?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
?? Might want to mention it in related work above, but the recent attention on adversarial examples/tricking models needs to be looked into for RL. Poisoning attacks relevant for online techniques: old bounds exist \textcite{DBLP:journals/jmlr/KloftL10}, new stuff concerns collaborative learners \cite{DBLP:conf/acsac/ShenTS16}, nothing for rl. Hot topic in deep networks \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18}, but naturally still relevant with even linear approximations or exact tabular case due to limits of the PAC assumption. There is now examination of evasion attacks wrt.\ RL \cite{DBLP:journals/corr/HuangPGDA17}!
?? evasion attacks by \textcite{DBLP:conf/sp/Carlini017}---all of these are computed by way of a general stochastic optimiser, such as \emph{Adam} \cite{DBLP:journals/corr/KingmaB14}. possible to apply something similar to our learned model to assess its security? would the suggested states even be valid? (i.e. since they're monotonically increasing for the most part).

?? More future work --- share knowledge between agents. ``Knowledge bases'' for this purpose? (see: Qianru).

@@

It all ends here...

?? Please write me once everything else is in place.

?? Overarching goal: works which \emph{respect the complexity of the network environment} (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?

We stress once again that this work still trails behind existing (exact) DDoS flow detection mechanisms, yet ...

>>>>

\emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).
We examine how the paradigm has been applied to the task of DDoS prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.

These flaws, poor handling of TCP traffic and high host density at egress points, catalyse and direct an exploration of the design space of RL models.
We improve the state-of-the-art through two agent classes designed to act on a per-flow basis, supported by an in-depth investigation of feature suitability and empirical evaluation.

\end{abstract}

@@

To date, there have been few applications of this class of algorithms towards intrusion prevention which make use of their full potential for online control, rather than using them as the basis of a classifier.
We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
In service of this goal, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention, improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

\section{Acting On Individual Sources}\label{sec:acting-on-individual-sources}
Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per source.
However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our enhancements to the RL model of \textcite{DBLP:journals/eaai/MalialisK15}, present a new model which draws on domain knowledge leveraged by SPIFFY \cite{DBLP:conf/ndss/KangGS16}, and offer an empirical investigation into the effectiveness of flow-level statistics on live decision-making.

@@

%?? Why does SPF only do well sometimes? Model is actually more difficult to learn, so it seems to do best when it has a larger set of decisions to learn from. But, it does worse for TCP?
\fakepara{Model performance}
Of the results presented, SPF's meandering performance is interesting given its far smaller action space.
It's natural to expect that this would the model easier to learn, but the additional state required appears to make the model \emph{harder}, beyond even the gains we might expect from choosing a non-zero discount factor.
Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
Paradoxically, Marl++ generally achieves the best performance yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.

@@

\section{Future Work}
%Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
%?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.

%?? IDEA: try out average reward, TD($\lambda$) methods as future work...
A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation.

%?? What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge? I think a good candidate is the sum of up and down throughputs (normalised by capacity sum), so long as \emph{neither exceeds the link capacity}. We can extend the team-based formulations similarly. This, in theory, promotes traffic diversity since it's not like flooding-based DDoS attacks are going to submit meaningful work to a server. The intuition, I suppose, is that certain classes of flow will have a small footprint in one direction which causes a sizeable increase in the other! Alternatively, monitor the health of canary flows which cross the team boundary (i.e. only one in-out link).
A problem we raised (without a clear solution) was the design of reward functions which do not rely upon heuristic estimates or a priori knowledge of benign traffic content.
If true online learning is desired (i.e., coping with a non-stationary environment), then such reward functions are sorely needed.
While $\bload{\cdot}$ is likely to be a good candidate for many deployments, we believe that internal ``canary flows'' (or other active measurement) will be effective.

%?? Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e. viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance. It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques. THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
Given that one of the advantages that RL methods can offer is the ability to handle non-stationary problems, it is important to propose and test sensible simulations or captures of evolving networks against these methods.
These scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.

%?? IDEA: Apply these techniques to programmable data planes etc. While it's pretty neat that what we have works assuming that ach router is a software (x86) switch running OVS, what might we need to consider when applying this to `real' switches? ``PDP can allow this to be added to real routers to make it efficient to keep \& process state in the manner we require, as well as enabling more adaptive deployment''. Cite P4, BPFabric, other work on PDPs?
%?? What concessions will we have to make in order to make per-flow processing more viable? Intelligent sampling/reanalysis of flows when needed (i.e. an external heuristic guiding method)? In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!
%?? More future work --- share knowledge between agents. ``Knowledge bases'' for this purpose? (see: Qianru).
Effective real-world deployment of RL-based defences cannot assume that switches in a network will support a custom version of OVS or other arbitrary software, introducing the question of whether agent training, execution and distribution may when using \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
We also expect it will be fruitful to look into \emph{how} agents may share knowledge with one another, such as via \emph{knowledge bases} or some other efficient communication mechanism.

%?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
%?? Might want to mention it in related work above, but the recent attention on adversarial examples/tricking models needs to be looked into for RL. Poisoning attacks relevant for online techniques: old bounds exist \textcite{DBLP:journals/jmlr/KloftL10}, new stuff concerns collaborative learners \cite{DBLP:conf/acsac/ShenTS16}, nothing for rl. Hot topic in deep networks \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18}, but naturally still relevant with even linear approximations or exact tabular case due to limits of the PAC assumption. There is now examination of evasion attacks wrt.\ RL \cite{DBLP:journals/corr/HuangPGDA17}!
%?? evasion attacks by \textcite{DBLP:conf/sp/Carlini017}---all of these are computed by way of a general stochastic optimiser, such as \emph{Adam} \cite{DBLP:journals/corr/KingmaB14}. possible to apply something similar to our learned model to assess its security? would the suggested states even be valid? (i.e. since they're monotonically increasing for the most part).
Although we believe that the security landscape for classical RL models is not \emph{identical} to that of neural-network based approaches (particularly with such noisy, volatile, and hard-to-control data), there is still immense value in determining the exact capabilities of a sufficiently powerful adversary as the risk of external control still exists.
In particular, we believe that poisoning attacks and evasion attacks merit close consideration.

@@

Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.

We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
Our evaluation shows that our new agent designs advance the state-of-the-art in RL-based DDoS prevention.

We stress once again that this work still trails behind the performance of exact DDoS flow detection mechanisms, yet we hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
%(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?
